---
title: "Interactive gene models with gggenomes and ggiraph (experimental)"
output:
  html_document:
    self_contained: true
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

This page demonstrates `geom_gene_interactive()`, an **experimental** drop-in
replacement for [`geom_gene()`](https://thackl.github.io/gggenomes/reference/geom_gene.html)
that renders gene models as interactive SVG elements via the
[ggiraph](https://davidgohel.github.io/ggiraph/) package. Hover over a gene
below to see its description/name/ID as a tooltip. All genes with the same name/description/ID
are also highlighted.

The example reuses the six-viral-genome comparison from the
[gggenomes README](https://github.com/thackl/gggenomes#readme). The only
change from the static version is swapping `geom_gene()` for
`geom_gene_interactive()` and mapping `tooltip` (and `data_id`, for highlighting)
to a gene identifier.

## Installation

`geom_gene_interactive()` is not yet part of a released version of
gggenomes. It currently lives on the `ggiraph-prototype` branch of
[thackl/gggenomes](https://github.com/thackl/gggenomes/tree/ggiraph-prototype)
on GitHub. Install it (together with ggiraph) with:

```r
# install.packages("remotes")
remotes::install_github("thackl/gggenomes", ref = "ggiraph-prototype")
install.packages("ggiraph")
install.packages("Hmisc")
```

```{r plot}
library(gggenomes)

p <- gggenomes(
  genes = emale_genes, seqs = emale_seqs, links = emale_ava,
  feats = list(emale_tirs, ngaros = emale_ngaros, gc = emale_gc)
) |>
  add_sublinks(emale_prot_ava) |>
  sync() + # synchronize genome directions based on links
  geom_feat(position = "identity", size = 6) +
  geom_seq() +
  geom_link(data = links(2)) +
  geom_bin_label() +
  geom_gene_interactive(aes(
    fill = name,
    tooltip = dplyr::coalesce(Note, name, feat_id),
    data_id = dplyr::coalesce(name, Note, feat_id)
  )) +
  geom_gene_tag(aes(label = name), nudge_y = 0.1, check_overlap = TRUE) +
  geom_feat(data = feats(ngaros), alpha = .3, size = 10, position = "identity") +
  geom_feat_note(aes(label = "Ngaro-transposon"),
    data = feats(ngaros),
    nudge_y = .1, vjust = 0
  ) +
  geom_wiggle(aes(z = score, linetype = "GC-content"), feats(gc),
    fill = "lavenderblush4", position = position_nudge(y = -.2), height = .2
  ) +
  scale_fill_brewer("Genes", palette = "Dark2", na.value = "cornsilk3")
```

```{r girafe, fig.width=9, fig.height=4.5}
ggiraph::girafe(
  ggobj = p,
  options = list(
    ggiraph::opts_sizing(rescale = TRUE),
    ggiraph::opts_zoom(min = 1, max = 5),
    ggiraph::opts_tooltip(opacity = .9),
    ggiraph::opts_hover(css = "stroke:black;stroke-width:2px;")
  )
)
```
